---
title: "Class06"
author: "Emily Darin"
format: pdf
---

written function for grading student homework 

```{r}
#' Title
#'
#' @param X 
#'
#' @return
#' @export
#'
#' @examples
  grade <- function(X){
    
    X[is.na(X)]<- 0 #makes NA = to 0 
    mean(X[-which.min(X)]) # calculated the mean after dropping the lowest score for each student 
    
  }

```


```{r}
gradebook<- read.csv("http://tinyurl.com/gradeinput", row.names=1)
gradebook
```

Using my function, I converted NA to 0, dropped the lowest score for each student and then averaged grades across students and dropped NA grades. 

```{r}
apply(gradebook,1,grade)
```

>Q2. Using your grade() function and the supplied gradebook, Who is the top scoring student overall in the gradebook? [3pts]

I used my function to calculate the results and then I asked R which student did the best by doing "which.max", this is student 18. 

```{r}
results <- apply(gradebook,1,grade)
which.max(results)
```

>Q3. From your analysis of the gradebook, which homework was toughest on students (i.e. obtained the lowest scores overall? [2pts]

The answer is Homework 2, to get this answer I took the median of each homework column and then asked R which median was the lowest. This gives me a better distribution of how students did than if I used the average. (mean is effected by outliers)

```{r}
homeworkmedian<-apply(gradebook,2,median,na.rm=TRUE)
which.min(homeworkmedian)
```

>Q4. Optional Extension: From your analysis of the gradebook, which homework was most predictive of overall score (i.e. highest correlation with average grade score)? [1pt]

Homework 5 is the most correlated to student average score. 

```{r}
gradebook[is.na(gradebook) ] <-0
cor(results, gradebook$hw1)

apply(gradebook,2,cor, x=results) 
    
      
```

