---
title: "Class10_ML"
author: "Emily Darin"
format: pdf

---
#Importing the candy data 


```{r}
candy_file <- "candy-data.csv"
candy <- read.csv(candy_file , row.names=1)
head(candy)
```

>Q1. How many different candy types are in this dataset?

13

```{r}
ncol(candy)
```

>Q2. How many fruity candy types are in the dataset?

38

```{r}
sum(candy$fruity)
```

## Data exploration 

Skimr package for a first peek at the dataset

```{r}
skimr::skim(candy)
```

>Q3. What is your favorite candy in the dataset and what is it’s winpercent value?

Twix and the winpercent is 81.6% 

```{r}
candy["Twix",]
```


>Q4. What is the winpercent value for “Kit Kat”?

The winpercent for kit kat is 76.7%

```{r}
candy["Kit Kat",]
```


>Q5. What is the winpercent value for “Tootsie Roll Snack Bars”?

The win percent for Tootsie Roll Snack Bars is 76%

```{r}
candy["Tootsie Roll Snack Bars",]
```


```{r}
x<-c(5,1,2,4,3)
order(x)

inds<- order(candy$winpercent)
head(candy[inds,])
tail(candy[inds,])



```

>Q6. Is there any variable/column that looks to be on a different scale to the majority of the other columns in the dataset?

The price percent of Twix is quite high compared to others. Reeses miniatures seem to have almost no sugar. 

>Q7. What do you think a zero and one represent for the candy$chocolate column?

It's binary for TRUE/FALSE 

>Q8. Plot a histogram of winpercent values

```{r}
library(ggplot2)

ggplot(candy, aes(winpercent,))+
geom_histogram(binwidth=5, fill="lightblue",color="black")+
xlab("Distribution of Winning Percentages")+
ylab("Count")+
theme_dark()
```

>Q9. Is the distribution of winpercent values symmetrical?

No, the distribution would be left skewed 

>Q10. Is the center of the distribution above or below 50%?

The center of the distribution is below 50% when the binwidth=8

>Q11. On average is chocolate candy higher or lower ranked than fruit candy?

Steps to answer: 
1. Find all chocolate candy (subset the candy table to be just chocolate and find the winpercent)
2. Summarize these wins into one metric (mean and standard deviation)
3. Repeat for fruit

```{r}

chocolate<-subset(candy$winpercent, candy$chocolate == 1)
chocolate

fruit<-subset(candy$winpercent, candy$fruity == 1)
fruit

t.test(chocolate,fruit)

```


>Q12. Is this difference statistically significant?

Yes these are both significantly different (p value of 2.871e-08)


####Graphing

>Q15 Make a first barplot of candy ranking based on winpercent values

```{r}
ggplot(candy)+
  aes(winpercent,row.names(candy))+
  ylab("Candy")+
  xlab("Win Percentages")+
  geom_col()

ggsave("barplot1.png", height=10,width=7)
```




>Q16. This is quite ugly, use the reorder() function to get the bars sorted by winpercent?


```{r}
ggplot(candy)+
  aes(winpercent, reorder(row.names(candy),winpercent))+
  ylab("Candy")+
  xlab("Win Percentages")+
  geom_col()

ggsave("barplot1.png", height=10,width=7)

```




```{r}

my_cols=rep("black", nrow(candy))
my_cols[as.logical(candy$chocolate)] = "chocolate"
my_cols[as.logical(candy$bar)] = "brown"
my_cols[as.logical(candy$fruity)] = "pink"
```

```{r}
ggplot(candy) + 
  aes(winpercent, reorder(rownames(candy),winpercent)) +
  geom_col(fill=my_cols) 

ggsave("barplot2.png", height=10,width=7)

```



> Q17. What is the worst ranked chocolate candy?

Tootsie smarties 


> Q18. What is the best ranked fruity candy?

Starburst

#looking at price percent 


```{r}
library(ggrepel)

# How about a plot of price vs win
ggplot(candy) +
  aes(winpercent, pricepercent, label=rownames(candy)) +
  geom_point(col=my_cols) + 
  geom_text_repel(col=my_cols, size=3.3, max.overlaps = 5)
```

>Q19. Which candy type is the highest ranked in terms of winpercent for the least money - i.e. offers the most bang for your buck?

Reeses minis 

>Q20. What are the top 5 most expensive candy types in the dataset and of these which is the least popular?

NLK,Ring pop, Nestle smarties, Hershey's Krackel, Hershey's Milk Chocolate and NLK is the least popular

```{r}
ord <- order(candy$pricepercent, decreasing = TRUE)
head( candy[ord,c(11,12)], n=5 )
```


```{r}
library(corrplot)

cij <- cor(candy)
corrplot(cij)
```
>Q22. Examining this plot what two variables are anti-correlated (i.e. have minus values)?

chocolate nougat 


>Q23. Similarly, what two variables are most positively correlated?

Chocolate and fruit 

#PCR

```{r}
pca <- prcomp(candy, scale=TRUE)
summary(pca)
```

```{r}
plot(pca$x[,1:2], col=my_cols, pch=16)
```

```{r}
my_data <- cbind(candy, pca$x[,1:3])
p <- ggplot(my_data) + 
        aes(x=PC1, y=PC2, 
            size=winpercent/100,  
            text=rownames(my_data),
            label=rownames(my_data)) +
        geom_point(col=my_cols)

p
```

```{r}
library(ggrepel)

p + geom_text_repel(size=3.3, col=my_cols, max.overlaps = 7)  + 
  theme(legend.position = "none") +
  labs(title="Halloween Candy PCA Space",
       subtitle="Colored by type: chocolate bar (dark brown), chocolate other (light brown), fruity (red), other (black)",
       caption="Data from 538")

```


```{r}


```
```{r}
par(mar=c(8,4,2,2))
barplot(pca$rotation[,1], las=2, ylab="PC1 Contribution")
```
>Q24. What original variables are picked up strongly by PC1 in the positive direction? Do these make sense to you?

Fruity, Hard and Pluribus this makes sense because hard fruit candies that come in packs of multiples. 