---
title: "Covid19"
format: pdf
---

```{r}
library(ggplot2)

vax <- read.csv("d4ff0cd9-d6d7-47a7-a5cb-28572c552298.csv")
head(vax)

```
>Q1. What column details the total number of people fully vaccinated? 

persons fully vaccinated

>Q2. What column details the Zip code tabulation area? 

Zip code tabulation data 

>Q3. What is the earliest date in this dataset? 

01-2022


>Q4. What is the latest date in this dataset? 

01-2021

```{r}
skimr::skim(vax)
```

>Q5. How many numeric columns are in this dataset?
9

>Q6. Note that there are “missing values” in the dataset. How many NA values there in the persons_fully_vaccinated column? 
4

Q7. What percent of persons_fully_vaccinated values are missing (to 2 significant figures)? 


```{r}
library(lubridate)
today()
vax$as_of_date <- ymd(vax$as_of_date)
today() - vax$as_of_date[1]
vax$as_of_date[nrow(vax)] - vax$as_of_date[1]


```


>Q9. How many days have passed since the last update of the dataset? 

546 days

>Q10. How many unique dates are in the dataset (i.e. how many different dates are detailed)? 

1160